---
title: "Forecasting the 2022 Governor Elections"
output: 
 github_document
date: "`r format(Sys.time(), '%B %d, %Y')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = TRUE)
rm(list = ls(all.names = TRUE)) 
```

# Settings

```{r message=FALSE}
library(tidyverse, quietly = TRUE)
library(ggdist, quietly = TRUE)
library(jsonlite, quietly = TRUE)
library(magrittr, quietly = TRUE)
library(boot, quietly = TRUE)
library(ggpubr, quietly = TRUE)

library(readxl, quietly = TRUE) # read data
library(lubridate, quietly = TRUE) # date
library(parsedate, quietly = TRUE)
library(readr, quietly = TRUE) # csv    
library(usdata, quietly = TRUE) # state names
library(RColorBrewer, quietly = TRUE)
library(rlang, quietly = TRUE)

library(parallel, quietly = TRUE)
library(rstan, quietly = TRUE)
library(tidybayes, quietly = TRUE)
library(cmdstanr, quietly = TRUE)
library(bayesplot, quietly = TRUE)
library(vecsets, quietly = TRUE)

FLAG_RUN_MODEL <- T

options(max.print = 10000) 
```


```{r message=FALSE}

date <- Sys.Date()

election_day <- ymd('2022-11-08')  

target_day <- election_day #min(date, election_day)

path <- paste0('../../EVA-midterm/midterm-web/public/forecasts/', date)

d_fte <- read_csv('data/fivethirtyeight-9.1.csv') %>% select(state, rep.name, dem.name)

dir.create(path)

```

# Theme set

```{r}
theme_set(theme_ggdist() + 
          theme(strip.background = element_blank(),
                panel.spacing = unit(2, "lines"),
                plot.title = element_text(hjust = .5)))

set.seed(1001)

```

```{r}

all_states <- c("AK", "AL", "AR", "AZ", "CA", "CO", "CT", "DC", "DE", "FL", "GA", "HI",
            "IA", "ID", "IL", "IN", "KS", "KY", "LA", "MA", "MD", "ME", "MI", "MN", "MO", "MS", "MT",
            "NC", "ND", "NE", "NH", "NJ", "NM", "NV", "NY", "OH", "OK", "OR", "PA", "RI", "SC", "SD", 
            "TN", "TX", "UT", "VA", "VT", "WA", "WI", "WV", "WY")
```

# Load data from fivethirtyeight

```{r}

today <- Sys.Date()

current_polls <- read_csv('https://projects.fivethirtyeight.com/polls/data/governor_polls.csv')

# Ratings from Cook report
current_status <- read_xlsx('./data/June_29.xlsx', sheet = 'June_29_states') %>% 
  rename(state_full = State) %>% 
  filter(state_full != 'Northern Mariana Islands') %>% 
  mutate(state = state2abbr(state_full))

current_rate <- read_xlsx('./data/June_29.xlsx', sheet = 'June_29_rate') %>% 
  rename(state_full = State) %>% 
  mutate(state = state2abbr(state_full))

head(current_polls)
head(current_status)
head(current_rate)
head(current_polls %>% 
  mutate(t = ymd(parse_date(start_date))))
```

## data clean

```{r}

df_polls_0 <- current_polls %>%
  filter(cycle == 2022 &
         stage == 'general') %>% 
  rename(state_full = state) %>%
  mutate(
         pct = pct / 100,
         # state name to abbr
         state = state2abbr(state_full),
         # clean up date
         start_date = mdy(start_date),
         end_date =  mdy(end_date)) %>% 
  filter(start_date >= ymd('2020-12-01'))


min_date = min(df_polls_0$start_date)

df_polls_1 <- df_polls_0 %>%
         mutate(date = round((start_date  - min_date + end_date  - min_date) / 2),
         # clean up
         methodology =  replace_na(methodology, 'unknown'),
         fte_grade =  replace_na(fte_grade, 'unknown')) %>% 
  dplyr::select(
    date,
    state,
    answer,
    candidate_name, 
    party,
    sample_size,
    pct,
    poll_id,
    pollster_id,
    question_id,
    pollster,
    display_name,
    fte_grade,
    methodology,
    population,
    state_full,
    start_date,
    end_date,
    url
  )



max_date = today

head(df_polls_1, n = 10)

# the number of states we can forecast
length(unique(df_polls_1$state))
```



```{r}
# transform cook report to get incumbency and the last election dem pct
current_status_short <-
  current_status %>% 
  rename(party = Party) %>% 
  mutate(incumbent_party = recode(party, 'Democratic' = 'DEM', 'Republican' = 'REP', 'DFL' = 'DFL')) %>% 
  mutate(last_dem = last_dem / (last_dem + last_rep) ) 
  
# attach cook report to polling data  
df_polls <-
  df_polls_1 %>% 
  left_join(current_status_short %>% dplyr::select(state, Status, last_dem, incumbent_party), by = 'state') %>% 
  left_join(current_rate %>% dplyr::select(state, Cook), by = 'state') %>% 
  rename(incumbent_status = Status) %>% 
  filter(!is.na(state))


```


Let's first get a sense of the polls

```{r fig.height=15, fig.width=15}

df_polls %>% 
  ggplot(aes(x = start_date, y = pct)) +
  geom_point(aes(color = party)) + 
  scale_y_continuous(limits = c(0, 1)) + 
  scale_x_date(date_labels = '%y-%m', limits = c(as.Date(min_date), today)) + 
  geom_label(x = today, y = .8, aes(label = Cook, fill = Cook), size = 2.5) +
  facet_wrap(. ~ state, ncol = 5, scales = 'free') 

```


## clean up and keep two party parts

We assume that all Democrat candidates' ballots will go to the final Democrat candidate, and all republicans' tickets will go to the final Republican candidate. This is an assumption that may not always be true.

We also discard other party candidates and their ballots. In some states like Alaska and New Mexicon, the third-party candidate is quite compelling.

A poll may have multiple questions. We assume that each question is a small poll.

```{r}
df_polls_two_party <- df_polls %>% 
  # These questions ask IND
  filter(question_id != 148891 & question_id != 162710) %>%
  filter(state != 'AK') %>% 
  group_by(question_id, party) %>% 
  summarise(grouped_pct = sum(pct), 
    date,
    state,
    sample_size,
    poll_id,
    pollster_id,
    question_id,
    pollster,
    display_name,
    fte_grade,
    methodology,
    population,
    state_full,
    start_date,
    incumbent_status,
    incumbent_party,
    last_dem,
    url
  ) %>%  
  distinct() %>% 
  spread(
    key = party, 
    value = grouped_pct
  ) %>% 
  mutate(
         n_dem = DEM * sample_size,
         n_rep = REP * sample_size,
         n_two_party =  round(n_dem + n_rep),
         dem_pct = n_dem / n_two_party,
         rep_pct = 1 - dem_pct,
         dem_num = round(dem_pct * n_two_party),
         date_index = round(as.numeric(date)),
         question_id = as.factor(question_id),
         poll_id = as.factor(poll_id),
         pollster_id = as.factor(pollster_id),
         last_dem_logit = logit(last_dem),
         methodology_recode = recode(
           methodology,
           "Online" = "Online",
           "Live Phone" = "Live Phone",
           "IVR" = "IVR",
           "unknown" = "unknown", 
           .default = "mixed"
         ),
         methodology_recode = as.factor(methodology_recode),
         population = as.factor(population)
         ) %>% 
  dplyr::select(n_two_party, 
                  dem_pct, rep_pct, 
                  dem_num, date_index, 
                  question_id, 
                  poll_id, 
                  pollster_id, 
                  pollster,
                  display_name,
                  methodology_recode,
                  population,
                  date, 
                  start_date,
                  state,
                  url,
                  incumbent_party,
                  last_dem) %>% 
  arrange(date) 


states <- unique(df_polls_two_party$state)

# check if there is NA
if(nrow(na.omit(df_polls_two_party)) != nrow(df_polls_two_party)) quit(save="ask")
```

## look at the data after clean-up

```{r fig.height=10, fig.width=15, warning=FALSE}
df_polls_two_party %>% 
  ggplot() +
  geom_line(mapping = aes(x = start_date, y = dem_pct), color = 'steelblue') + 
  geom_point(mapping = aes(x = start_date, y = dem_pct), size = 1, color = 'steelblue') + 
  geom_line(mapping = aes(x = start_date, y = rep_pct), color = 'salmon') +
  geom_point(mapping = aes(x = start_date, y = rep_pct), size = 1, color = 'salmon') + 
  geom_hline(yintercept = .5, color = 'gray', linetype = 'dashed') + 
  scale_y_continuous(limits = c(0, 1)) + 
  scale_x_date(date_labels = '%y-%m', limits = c(as.Date(min_date), today)) + 
  facet_wrap(. ~ state, ncol = 5, scales = 'free') 
```

Let's check how many polls we have for a state

```{r}
df_polls_two_party %>%  
  group_by(state) %>% 
  summarise(n())
```


# Modeling preparation




```{r include=FALSE}
MCMC_SETTINGS <- list(
  N_ITER = 15000,
  N_WARMUP = 5000,
  N_THIN = 2,
  N_CHAINS = 4,
  DELTA = 0.9,
  TREE_DEPTH = 15,
  SEED = 1234
)

rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
```


# Model 

## stan code

```{r}
m5_stan ="
data{
  int N_questions;   
  int N_polls;    
  int N_days;   
  int N_states;  
  int N_pollsters;    
  int N_population;
  int N_methods; 
  int direction_flag;
  
  array[N_questions] int<lower = 1, upper = N_states> state_index; // State index
  array[N_questions] int<lower = 1, upper = N_days>  day_state_index;   // Day index
  
  array[N_questions] int<lower = 1, upper = N_polls>  poll_index; // poll index
  array[N_questions] int<lower = 1, upper = N_pollsters> pollster_index; // which pollster
  array[N_questions] int<lower = 1, upper = N_methods> methods_index;   // which polling method
  array[N_questions] int<lower = 1, upper = N_population> population_index;   // which method

  vector[N_states] last_dem_logit;   // which incumbency party in that state?
 
  array[N_questions] int dem_num;
  array[N_questions] int n_two_party;
  
  real state_day_sigma_scaler;
  real states_poll_bias_scaler;
  real sigma_scaler_pollster;
  real sigma_scaler;
}

parameters {
  
  vector[N_pollsters] mu_pollster; 
  vector[N_methods] mu_methods; 
  vector[N_population] mu_population; 
  
  vector[N_polls] poll_noise;
  vector[N_states] polling_bias_states;  
  
  matrix[N_states, N_days] err;
  vector[N_states] raw_mu_first_day;

}

transformed parameters {
 

  matrix[N_states,N_days] mu; 

  // this is forward
  if(direction_flag > 0){
    
    mu[:, 1] = raw_mu_first_day;
    
    for (t in 2:N_days) {
      mu[:, t] = mu[:,  t - 1] + err[:, t] * state_day_sigma_scaler;
    }
    
  }else{
    // this is backward
    mu[:, N_days] = raw_mu_first_day;
    
    for (t in 1:(N_days-1)) {
      mu[:, N_days - t] = mu[:,  N_days - t + 1] + err[:, N_days - t + 1] * state_day_sigma_scaler;
    }
  }



 vector[N_questions] logit_dem_pct;  

 for (i in 1:N_questions){
   logit_dem_pct[i] =
    mu[state_index[i], day_state_index[i]] + 
    mu_pollster[pollster_index[i]] * sigma_scaler_pollster + 
   // mu_incumbent_party[incumbent_party_index[state_index[i]]] * sigma_scaler + 
    mu_methods[methods_index[i]] * sigma_scaler +
    mu_population[population_index[i]] * sigma_scaler + 
    polling_bias_states[state_index[i]] * states_poll_bias_scaler + 
    poll_noise[poll_index[i]] * sigma_scaler;
 }

}

model {

  mu_pollster ~ std_normal();
  //mu_incumbent_party ~ std_normal();
  mu_methods ~ std_normal();
  mu_population ~ std_normal();
  poll_noise ~  std_normal();
  polling_bias_states ~  std_normal();
  raw_mu_first_day ~ normal(last_dem_logit, .3); 
  
  to_vector(err) ~  std_normal();

  dem_num ~ binomial_logit(n_two_party, logit_dem_pct);
  
}

"

if(exists('out')) rm(out)
if(exists('df_prediction')) rm(df_prediction)
if(exists('prediction')) rm(prediction)
if(exists('df_mu_methods')) rm(df_mu_methods)
if(exists('df_poll_noise')) rm(df_poll_noise)

f <- write_stan_file(code = m5_stan, dir = 'stan', basename = 'm5.stan')
```

## transform data to pass to stan

```{r}
df_polls_two_party <-
  df_polls_two_party %>% 
  ungroup() %>% 
  mutate(state_index = as.numeric(as.factor(state)),
         day_state_index = as.integer(date) + 1,
         day = min_date + date, 
         pollster_index = as.numeric(as.factor(pollster_id)),
         poll_index = as.numeric(as.factor(poll_id)),
         methods_index = as.numeric(as.factor(methodology_recode)),
         population_index = as.numeric(as.factor(population)),
         incumbent_party_index = as.numeric(as.factor(incumbent_party))
         )

states <- levels(as.factor(df_polls_two_party$state))
```


## pass data to stan


```{r}

pulled_incumbent_party_index <- df_polls_two_party %>% 
                                dplyr::select(c(state_index, incumbent_party_index)) %>% 
                                arrange(state_index) %>% distinct() %>% pull(incumbent_party_index)


pulled_last_dem_logit <- logit(df_polls_two_party %>% 
                                dplyr::select(c(state_index, last_dem)) %>% 
                                arrange(state_index) %>% distinct() %>% pull(last_dem))

sigma_scaler = 0.04 
sigma_scaler_pollster = 0.06

state_day_sigma_scaler = 0.016
# 0.0165 from the economist's 2020 presidential election
# 0.032 if we want to double it

states_poll_bias_scaler = .076
# .076 result in +- in total 0.07585401
# 0.15 if we want to double it
# 0.2 if we want to make 20

# we use the forward model
direction_flag = 1

m5_data = list( 
  sigma_scaler = sigma_scaler,
  sigma_scaler_pollster = sigma_scaler_pollster,
  state_day_sigma_scaler = state_day_sigma_scaler, 
  states_poll_bias_scaler = states_poll_bias_scaler,
  direction_flag = direction_flag,
  # states_sd_scaler = states_sigma_scaler,
  # states_sd_scaler = sd_rough$sd_scaler,
  # states_avg_sigma = states_avg_sigma,
  N_questions = length(unique(df_polls_two_party$question_id)),
  N_polls = length(unique(df_polls_two_party$poll_index)),
  N_days = as.integer(round(difftime(election_day, min_date))) + 1,
  N_states = length(unique(df_polls_two_party$state_index)),
  N_pollsters = length(unique(df_polls_two_party$pollster_index)),
  N_population = length(unique(df_polls_two_party$population_index)),
  N_methods = length(unique(df_polls_two_party$methods_index)),
  N_incumbent_party = length(unique(df_polls_two_party$incumbent_party)),
    
  state_index = df_polls_two_party$state_index,
  day_state_index = df_polls_two_party$day_state_index,
  pollster_index = df_polls_two_party$pollster_index,
  poll_index = df_polls_two_party$poll_index,
  methods_index = df_polls_two_party$methods_index,
  population_index = df_polls_two_party$population_index,
  incumbent_party_index = pulled_incumbent_party_index,
    
  dem_num = df_polls_two_party$dem_num,
  n_two_party = df_polls_two_party$n_two_party,
  last_dem_logit = pulled_last_dem_logit
)
  

```

## model fit


```{r}

if (FLAG_RUN_MODEL) {
  
   if (file.exists("stan/m5")) {
      file.remove("stan/m5")
   }
  
  m5 <- cmdstanr::cmdstan_model('stan/m5.stan',
                            compile = TRUE,
                            force = TRUE
                            #force_recompile = TRUE,
                            #https://rdrr.io/github/stan-dev/cmdstanr/src/R/model.R
                            #stanc_options = list("auto-format")
                           )
  
  print(m5)
  #print(m5_data)
  
  fit <- m5$sample(
    data = m5_data,
    seed = 99,
    parallel_chains = MCMC_SETTINGS$N_CHAINS, 
    chains = MCMC_SETTINGS$N_CHAINS,
    iter_warmup = MCMC_SETTINGS$N_WARMUP,
    iter_sampling = MCMC_SETTINGS$N_WARMUP,
    refresh = MCMC_SETTINGS$N_WARMUP * .2,
    max_treedepth = MCMC_SETTINGS$TREE_DEPTH,
    adapt_delta = MCMC_SETTINGS$DELTA,
    init = 0
  )
  
  fit$cmdstan_diagnose()
  #fit$cmdstan_summary()
  

  out <- rstan::read_stan_csv(fit$output_files())
  
  
  # save model
  write_rds(out, sprintf('rds/out_%s.rds', date), compress = 'gz')
  
  write_rds(fit, sprintf('rds/fit_%s.rds', date), compress = 'gz')
  
  rm(fit)
  gc(full = T)
  
} else {
  out <- read_rds(sprintf('rds/out_%s.rds', date))
}
```


## posterior prediction


```{r}
prediction <-
  rstan::extract(out, pars = "mu")[[1]]
```

```{r}
nX <- MCMC_SETTINGS$N_CHAINS * MCMC_SETTINGS$N_WARMUP
nY <- m5_data$N_states
nZ <- m5_data$N_days

dim(prediction) <- nX * nY * nZ

df_prediction <- tibble(prediction = inv.logit(prediction), 
                   .draw = rep(1:nX, times = nY * nZ),
                   day_index = rep(1:nZ, each = nY * nX),
                   state_index = rep(rep(1:nY, each = nX), times = nZ),
                   day = min_date + day_index - 1,
                   state = states[state_index]
                 )

```


## current day

```{r}
prediction_summary <-
  df_prediction %>% 
    group_by(day, state) %>% 
    median_qi(prediction) %>% 
    mutate(diff = .upper - .lower)
```


Let's check the model prediction for the entire election season

```{r fig.height=15, fig.width=15}

prediction_summary %>% 
  ggplot() + 
    geom_lineribbon(
      aes(
      x = day,
      y = prediction,
      ymin = .lower,
      ymax = .upper
    ),
    fill = 'lightgray',
    size = 1
    ) +
  geom_point(
    df_polls_two_party,
    mapping = aes(x = day, y = dem_pct)
  ) + 
  geom_hline(df_polls_two_party %>% select(state, last_dem) %>% distinct(),
             mapping = aes(yintercept = last_dem), linetype = 'dotted', color = 'purple', size = .8) + 
  geom_vline(xintercept = today, linetype = 'dotted') + 
  geom_hline(yintercept = 0.5, linetype = 'dotted', size = .8) + 
  facet_wrap(state ~ ., ncol = 5, scale = 'free_x')
```

Get summary on election day

```{r fig.height=6, fig.width=5}
election_day_prediction <- df_prediction %>%
  filter(day == target_day) %>% 
  rename(dem = prediction) %>% 
  mutate(rep = 1 - dem) %>% 
  group_by(state) 

election_day_prediction %>% 
  mean_qi(dem, rep, .width = .95) 

election_summary <- 
  election_day_prediction %>% 
    mean_qi(dem, rep, .width = .8) %>%
    merge(election_day_prediction %>% 
        summarise(prob.dem = mean(dem > .5), prob.rep = mean(1 - dem > .5))) #%>%
```


```{r fig.height=6, fig.width=5, include=FALSE}
write_csv(
  election_day_prediction %>% 
    mean_qi(dem, rep, .width = c(.8, .95)) %>%
    mutate(state_day_sigma_scaler = state_day_sigma_scaler,
           states_poll_bias_scaler = states_poll_bias_scaler, 
           direction_flag = direction_flag), 
  file = paste0('model_comparison/', state_day_sigma_scaler, '__', states_poll_bias_scaler, '__', direction_flag, '.csv')
)
```


Show 80% credible intervals and probabilities (like fivethirtyeight)

```{r fig.height=6, fig.width=5}
election_summary %>% 
  rowwise() %>% 
  mutate(lower = dem.lower,
         lower.end = max(dem.lower, min(.5, dem.upper)),
         upper= dem.upper,
         upper.start = min(dem.upper, max(.5, dem.lower)),
         prob = max(prob.dem, prob.rep),
         diff_prob = abs(prob - .5)) %>%
  arrange(desc(diff_prob)) %>% 
  mutate(state = factor(state, unique(state))) %>% 
  ggplot() + 
  geom_segment(aes(x = lower, xend = lower.end, y = state, yend = state), size = 4, color = 'salmon') +
  geom_segment(aes(x = upper.start, xend = upper, y = state, yend = state), size = 4, color = 'steelblue4') +
  geom_vline(xintercept = c(.25, .375, .5, .625, .75), linetype='dotted') +
  geom_text(x = .8, aes(y = state, label = prob)) + 
  scale_x_continuous(limits = c(.1, .8), breaks = c(.25, .375, .5, .625, .75), labels = c('+50', '+25', '0', '+25', '+50'))

```


## check other terms

First, let's look at the polling bias for each state

```{r}
polling_bias_states <-
  rstan::extract(out, pars = "polling_bias_states")[[1]]

dim(polling_bias_states) <- nX * m5_data$N_states

df_polling_bias_states <- tibble(
         posterior = polling_bias_states,
         .draw = rep(1:nX, times =  m5_data$N_states),
         state_index = rep(1: m5_data$N_states, each =  nX),
         state = states[state_index]
       ) 

df_polling_bias_states %>% 
    group_by(state) %>% 
    ggplot(aes(x = state, y = posterior)) +
    stat_interval() +
    geom_hline(yintercept=c(0, -2, 2), linetype = 'dotted') + 
    scale_color_brewer() +
    scale_y_continuous()

# save memory
rm(polling_bias_states)
rm(df_polling_bias_states)
```


also pollster bias

```{r}
mu_pollster <-
  rstan::extract(out, pars = "mu_pollster")[[1]]

dim(mu_pollster) <- nX * m5_data$N_pollsters

df_mu_pollster <- tibble(
         posterior = mu_pollster,
         .draw = rep(1:nX, times =  m5_data$N_pollsters),
         pollster_id = paste0('pollster_', rep(1:m5_data$N_pollsters, each = nX))
       ) 

df_mu_pollster %>% 
    group_by(pollster_id) %>% 
    ggplot(aes(x = pollster_id, y = posterior)) +
    stat_interval() +
    geom_hline(yintercept=c(0, -2, 2), linetype = 'dotted') + 
    scale_color_brewer() +
    scale_y_continuous()

# save memory
rm(mu_pollster)
rm(df_mu_pollster)
```

Also different voter types

```{r}

mu_population <-
  rstan::extract(out, pars = "mu_population")[[1]]

dim(mu_population) <- nX * m5_data$N_population

df_mu_population <- tibble(
         posterior = mu_population,
         .draw = rep(1:nX, times =  m5_data$N_population),
         id = paste0('population_', rep(1:m5_data$N_population, each = nX))
       ) %>% 
    group_by(id) 

df_mu_population %>% 
    ggplot(aes(x = id, y = posterior)) +
    stat_interval() +
    geom_hline(yintercept=c(0, -2, 2), linetype = 'dotted') + 
    scale_color_brewer() +
    scale_y_continuous()

# save memory
rm(mu_population)
rm(df_mu_population)
```

The effects of different polling methods

```{r}

mu_methods <-
  rstan::extract(out, pars = "mu_methods")[[1]]

dim(mu_methods) <- nX * m5_data$N_methods

df_mu_methods <- tibble(
         posterior = mu_methods,
         .draw = rep(1:nX, times =  m5_data$N_methods),
         id = paste0('method_', rep(1:m5_data$N_methods, each = nX))
       ) 

df_mu_methods %>% 
    group_by(id) %>% 
    ggplot(aes(x = id, y = posterior)) +
    stat_interval() +
    geom_hline(yintercept=c(0, -2, 2), linetype = 'dotted') + 
    scale_color_brewer() +
    scale_y_continuous()

# save memory
rm(mu_methods)
rm(df_mu_methods)
```

Each individual poll's nosie

```{r}

poll_noise <-
  rstan::extract(out, pars = "poll_noise")[[1]]

m5_data$N_polls <- m5_data$N_polls 

dim(poll_noise) <- nX * m5_data$N_polls

df_poll_noise <-
tibble(
         posterior = poll_noise,
         .draw = rep(1:nX, times =  m5_data$N_polls),
         id = paste0('polls_', rep(1:m5_data$N_polls, each = nX))
       ) 

df_poll_noise %>% 
    group_by(id) %>% 
    ggplot(aes(x = id, y = posterior)) +
    stat_interval() +
    geom_hline(yintercept=c(0, -2, 2), linetype = 'dotted') + 
    scale_color_brewer() +
    scale_y_continuous()

# save memory
rm(poll_noise)
rm(df_poll_noise)
```


```{r  eval=FALSE}
# This was used and then found incumbent effect does not matter
mu_incumbent_party <-
  rstan::extract(out, pars = "mu_incumbent_party")[[1]]

dim(mu_incumbent_party) <- nX * m5_data$N_incumbent_party

tibble(
         posterior = mu_incumbent_party,
         .draw = rep(1:nX, times =  m5_data$N_incumbent_party),
         id = paste0('party_', rep(1:m5_data$N_incumbent_party, each = nX))
       ) %>% 
    group_by(id) %>% 
    ggplot(aes(x = id, y = posterior)) +
    stat_interval() +
    geom_hline(yintercept=c(0, -1, 1), linetype = 'dotted') + 
    xlab('incumbent_party') + 
    scale_color_brewer() +
    scale_y_continuous()
```




# To web

```{r}

write.csv(election_day_prediction %>% select(.draw, dem, rep, day, state), file = paste0('model-output/', date, '_election_day_prediction.csv'))

```

The following code is used only for generating data for the website. We omit explanations.

```{r include=FALSE}

posterior_dem_election_day <- df_prediction   %>%   
                    filter(day == target_day) %>%    
                    rename(dem = prediction)
```

```{r include=FALSE}
N_pdraws <- MCMC_SETTINGS$N_CHAINS * (MCMC_SETTINGS$N_WARMUP)
N_states = length(unique(df_polls_two_party$state))
N_draws <- 100
step <- N_pdraws / N_draws
bin_width <-   100 / N_draws
trailer <- bin_width / 2
bin_func <- function(d){
  return(floor(100 * d / bin_width) * bin_width + trailer)
}
```

## posterior samples

```{r include=FALSE}
posterior_draws_election_day <- posterior_dem_election_day %>%
  dplyr::select(.draw, state, dem) %>%
  group_by(state) %>% 
  mutate(
    rep = 1 - dem,
    nrank =  rank(dem, N_pdraws, ties.method = 'random'),
    outcome = if_else(dem - rep > 0, "dem",
                      if_else(dem - rep < 0, 'rep',  'tie')),
    outcome_index = case_when(outcome == 'dem' ~ 0,
                              outcome == 'rep' ~ 2,
                              outcome == 'tie' ~ 1),
    bin_dem = bin_func(dem),
    bin_rep = 100 - bin_func(dem),
  )

```

## ci, density, etc 

```{r include=FALSE}

day_state_summary <- posterior_draws_election_day %>% 
    group_by(state) %>% 
    mean_qi(dem, rep) %>% 
    dplyr::select(-c('.width', '.point', '.interval'))


density_bin_dem_election_day <-  posterior_draws_election_day %>% 
    mutate(bin_dem = as.factor(bin_dem)) %>% 
    group_by(state, bin_dem) %>% 
    summarise(n_bin = n(), density_bin = n_bin / N_pdraws) %>% 
    mutate(name = 'dem')


density_bin_rep_election_day <- posterior_draws_election_day %>% 
    mutate(bin_rep = as.factor(bin_rep)) %>% 
    group_by(state, bin_rep) %>% 
    summarise(n_bin = n(), density_bin = n_bin / N_pdraws) %>% 
    mutate(name = 'rep')
```

## prob

```{r include=FALSE}
prob <- posterior_draws_election_day %>% 
  group_by(outcome, state) %>% 
  summarise(prob = n() / N_pdraws) %>%
  pivot_wider(names_from = outcome, values_from = prob, names_prefix = 'prob.')

day_var <- posterior_draws_election_day   %>%
  group_by(state) %>%
  summarise(rep.var = var(rep * 100), 
            dem.var = var(dem * 100)) 


# merge into summary
day_state_summary <- merge(day_state_summary, prob, no.dups = TRUE)
day_state_summary <- merge(day_state_summary, day_var, no.dups = TRUE)

head(day_state_summary)
```


```{r include=FALSE}

no_race <- vsetdiff(all_states, unique(current_status$state))
                    
day_state_summary <- day_state_summary %>% 
  mutate(prob.dem = replace_na(prob.dem, 0),
         prob.rep = replace_na(prob.rep, 0)) %>%
  full_join(current_status, by = 'state') %>%
  mutate(has_election = TRUE) %>%
  add_row(state = no_race,
          has_election = FALSE) %>%
  arrange(abs(prob.dem - .5))
```



## save day summary

```{r include=FALSE}

str <- paste0(  
  '{',
  '\"date\":', '\"', date,'\"',',',
  '\"summary\":', toJSON(day_state_summary),
  '}'
)


str

write(str,  file = paste0(path, '/day-summary.json'))
```


## all days' summary


```{r include=FALSE}

full_state_summary_ci <- df_prediction   %>%
  rename(dem = prediction) %>%
  mutate(rep = 1 - dem) %>% 
  group_by(state, day) %>%
  mean_qi(rep, dem) %>%
  dplyr::select(-c('.width', '.point', '.interval'))
  

full_state_summary_prob <- df_prediction   %>%
  rename(dem = prediction) %>%
  mutate(rep = 1 - dem) %>% 
  group_by(state, day) %>%
  summarise(rep.prob = mean(rep > .5), dem.prob = mean(dem > .5), tie.prob = 1- dem.prob-rep.prob)
  


full_state_summary <- merge(full_state_summary_ci, full_state_summary_prob, no.dups = TRUE)
# full_state_summary <- merge(full_state_summary, full_state_summary_var, no.dups = TRUE)

head(full_state_summary)
```

## get poll info

```{r include=FALSE}
 df_polls_two_party
```

## sample 100 posterior draws

```{r fig.height=7, fig.width=10, include=FALSE}

draws <- posterior_draws_election_day %>%
  filter(((nrank + step / 2) %% step) == 0)


draws %>%
  ggplot() +
  geom_dots(aes(x = bin_dem)) +
  facet_wrap(state ~ .)

```

```{r include=FALSE}
# for plinko

generate_seq <- function(l, r, new_start){
  if(l %% 1 != 0 | r %% 1 != 0)   print(paste(l, r, new_start))
  original_seq <- sample(rep(c(-.5 * bin_width, .5 * bin_width), times=c(l,r))) 
  cum_seq <- c(new_start, cumsum(original_seq) + new_start)
  return(cum_seq)
}


trailer <- bin_width / 2

```

```{r include=FALSE}
for(s in states) {
  
  last_names <- d_fte %>% filter(state == s)
  
  # 100 samples
  samples <- draws %>%
    filter(state == s)  %>%
    mutate(across(everything(), as.character))
  
  sample_summary <- samples %>% 
    summarise(prob.dem = mean(outcome == 'dem'),
              prob.rep = mean(outcome == 'rep'),
              prob.tie = mean(outcome == 'tie'))
  
  # time plot
  time_state_summary <- full_state_summary %>%
    filter(state == s) %>%
    mutate(across(everything(), as.character))
  

  current_summary <- day_state_summary %>% 
    filter(state == s)
  
  # plinko
  start_dem <-
    floor(current_summary['dem'][[1]] * 100 / bin_width) * bin_width + trailer
  start_rep <-
    floor(current_summary['rep'][[1]] * 100 / bin_width) * bin_width  + trailer
  bin_n_dem <-
    ceiling(4 * current_summary['dem.var'][[1]] / bin_width ^ 2)
  bin_n_rep <-
    ceiling(4 * current_summary['rep.var'][[1]]  / bin_width ^ 2)
  max_density <-
    max(
      max(density_bin_rep_election_day$density_bin),
      max(density_bin_dem_election_day$density_bin)
    )
  
  plinko_height <- max(bin_n_dem, bin_n_rep)
  
  if (plinko_height %% 2 != 0)
    plinko_height = plinko_height + 1
  
  print(paste(plinko_height, s))
  
  plinko_steps <- draws %>%
    filter(state == s)  %>%
    mutate(
      diff_rep = bin_rep - start_rep,
      diff_dem = bin_dem - start_dem,
      l_rep =  .5 * (plinko_height - 2 * diff_rep / bin_width),
      r_rep =  .5 * (plinko_height + 2 * diff_rep / bin_width),
      l_dem =  .5 * (plinko_height - 2 * diff_dem / bin_width),
      r_dem =  .5 * (plinko_height + 2 * diff_dem / bin_width)
    ) %>%
    rowwise() %>%
    mutate(# steps_seq_dem = list(generate_seq(l_dem, r_dem, start_dem)),
      steps_seq = list(generate_seq(l_rep, r_rep, start_rep))) %>%
    select(.draw,
           state,
           dem,
           rep,
           outcome,
           outcome_index,
           bin_dem,
           bin_rep,
           steps_seq) %>%
    mutate(
      .draw = as.character(.draw),
      dem = as.character(dem),
      rep = as.character(rep),
      outcome = as.character(outcome),
      outcome_index = as.character(outcome_index),
      bin_dem = as.character(bin_dem),
      bin_rep = as.character(bin_rep)
    ) %>%
    ungroup()
  
  
  # overall intervals stuff
  state_summary <- day_state_summary %>%
    filter(state == s) %>%
    filter() %>% 
    mutate(winner = if_else(prob.dem > .5, 'dem', 'rep'))
  
  # polls
  state_polls <- df_polls_two_party %>%
    arrange(desc(date)) %>%
    dplyr::select(-date) %>%
    filter(state == s)
  
  # for density   
  day_boundraies <- time_state_summary %>% 
    filter(day == target_day & state == s) %>% 
    mutate(rep = bin_func(as.numeric(rep)), 
           dem = bin_func(as.numeric(dem)),
           rep.lower = bin_func(as.numeric(rep.lower)),
           rep.upper = bin_func(as.numeric(rep.upper)),
           dem.lower = bin_func(as.numeric(dem.lower)),
           dem.upper = bin_func(as.numeric(dem.upper)))
  
  str <- paste0(
    '{',
    '\"date\":',
    '\"',
    date,
    '\"',
    ',',
    '\"state\":',
    '\"',
    s,
    '\"',
    ',',
    '\"last_names\":',
    toJSON(last_names),
    ',',
    '\"samples\":',
    toJSON(sample(plinko_steps)),
    ',',
    '\"sample_summary\":',
    toJSON(sample_summary),
    ',',
    '\"state_summary\":',
    toJSON(state_summary),
    ',',
    '\"time_summary\":',
    toJSON(time_state_summary),
    ',',
    '\"polls\":',
    toJSON(state_polls),
    ',',
    '\"bin_width\":\"',
    bin_width,
    '\"',
    ',',
    '\"plinko_height\": \"',
    plinko_height,
    '\",',
    '\"plinko_center\": \"',
    start_rep,
    '\"',
    ',',
    # intervals
    '\"density_bin_rep_today\":',
    toJSON(density_bin_rep_election_day %>%
             filter(state == s) %>% 
             filter(as.numeric(as.character(bin_rep)) >= day_boundraies$rep.lower[1] & 
                    as.numeric(as.character(bin_rep)) <= day_boundraies$rep.upper[1]) %>% 
             mutate(across(
               everything(), as.character
             ))) ,
    ',',
    '\"density_bin_dem_today\":',
    toJSON(density_bin_dem_election_day %>%
             filter(state == s) %>% 
             filter(as.numeric(as.character(bin_dem)) >= day_boundraies$dem.lower[1] & 
                    as.numeric(as.character(bin_dem)) <= day_boundraies$dem.upper[1]) %>% 
             mutate(across(
               everything(), as.character
             )))
    ,
    '}'
  )
  
  write(str,  file = paste0(path, '/', s, '.json'))
  
  str
}

```






